Ab Initio Effective One-Electron Potential Operators: Applications for Charge-Transfer Energy in Effective Fragment Potentials

### 2019-11-29 papers/2_oep_technique/paper-r2.tex
Paper 2 added plotting sources

The apparent difficulty in theoretically characterising the CT energy
in terms of the interacting molecular fragments
is indeed a challenge in developing modern force fields 
or \emph{ab initio} fragmentation methods\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}
for modeling structure
and dynamics in condensed phases.
This needs to be contrasted with the Coulombic electrostatics,
non\hyp{}Coulombic repulsion (i.e., due to Pauli exclusion principle), 
dispersion and induction for instance, 
which to a certain extent
can be well described by, respectively, 
the multipole moments of charge density distributions,\cite{Sokalski.Poirier.CPL.1983,
Etchebest.Lavery.Pullman.TheorChimActa.1982,Stone.JCTC.2005}
the van der Waals repulsive and attractive potentials, 
as well as the polarizability models, i.e., all being relatively simple and computationally
inexpensive to evaluate mathematical models. 
Due to this reason, the CT effects 
%while usually of significant and non\hyp{}negligible magnitude as compared 
%to the total interaction energy, 
are not explicitly included in any molecular mechanics force field developed up to date.
%such as OPLS, Amber, CHARMM, Gromos or even polarizale AMOEBA force field. 
The only way to
incorporate the CT effects in the condensed\hyp{}phase simulations,
apart from performing full QM electronic structure simulations,
is through the second generation of the effective fragment potential 
(EFP2) method,\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013,
   Nguyen.Pachter.Day.JCP.2014,
   Day.Jensen.Gordon.Webb.Stevens.Krauss.Garmer.Basch.Cohen.JCP.1996}
developed by Gordon group.\cite{Li.Gordon.Jensen.JCP.2006,Xu.Gordon.JCP.2013}
EFP2 is an \emph{ab initio} force field
derived on the grounds of the first\hyp{}principles perturbation theory
at the HF level and including the intermolecular dispersion effects
by the response theory. 
%based on the imaginary frequency\hyp{}dependent dipole\hyp{}dipole polarizabilities.
That is to say, the total intermolecular 
interaction potential, which can also be perceived as an approximation
to SAPT at HF level with added intermolecular dispersion correction, is approximated as 
%
\begin{equation}\label{e:u-efp2}
 U^{\rm EFP2} \approx U^{\rm Coul} + U^{\rm Ex-Rep} + U^{\rm Ind} + U^{\rm Disp} + U^{\rm CT} \;.
\end{equation}
%
where $U^{\rm Coul}$ is the Coulombic interaction energy of the unperturbed charge\hyp{}density
distributions of the monomers, approximated by the distributed multipole approximation,
$U^{\rm Ex-Rep}$ is the exchange\hyp{}repulsion energy originating from the Pauli exclusion
principle, that could be roughly understood as the Heitler\hyp{}London interaction
energy with $U^{\rm Coul}$ subtracted, $U^{\rm Ind}$ and $U^{\rm Disp}$
are the induction and dispersion energies obtained from the distributed polarizability
approximation, and finally $U^{\rm CT}$ is the CT energy, the focus of this work.

@@

                       \frac{ \varphi^*_\eta({\bf r}_1) \hat{v} \phi_i({\bf r}_2) } 

@@

Four complexes: 
(i) (H$_2$O)$_2$, 
(ii) H$_2$O--CH$_3$OH, 
(iii) H$_2$O--NH$_4^+$ and 
(iv) NO$_3^-$--NH$_4^+$,
were chosen as model systems to analyse the asymptotic dependence 
of CT/HF0 energy. 
The reference (zero\hyp{}displacement) geometries
were obtained by
energy\hyp{}optimizations at the HF/6-31+G(d,p) level,
as implemented in 
the {\sc Gaussian~16} quantum chemistry program package.\cite{Gaussian16}
Subsequently, 30 displaced geometries for each model complex
were obtained by translating one of the monomers along the vector 
co\hyp{}linear with the H-bond or N--N distance in the case of ammonium nitrate.
The reference structures as well as the translation vectors are
indicated in the insets of Figure~\ref{f:fig-1} and S1.
To perform statistical error analysis, structural databases
of bi\hyp{}molecular complexes in the non\hyp{}covalent
interactions database NCB31 developed by the Truhlar's 
group,\cite{Zhao.Schultz.Truhlar.JCTC.2006,
Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
as well as the BioFragment Database subset BBI for backbone\hyp{}backbone
interactions in proteins,\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017} 
as implemented in the {\sc Psi4}
program,\cite{Psi4.JCTC.2017}
were utilized.
In particular, the subsets from the NCB31 database were separately considered:
the HB6/04 hydrogen bonding database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
the DI6/04 dipoleinteraction database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
the CT7/04 charge-transfer complex database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
the WI7/05 weak interaction database,\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
and PPS5/05 the $\pi$-$\pi$ stacking database.\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}

@@

%
\begin{figure*}[h]
\includegraphics[width=0.9\textwidth]{fig-2.pdf}
\caption{\label{f:fig-2} {\bf Performance of the OEP and EFP2 methods for the CT/HF0 charge transfer interaction energy
in small bi\hyp{}molecular neutral complexes.}
(a) accuracy of the ERI elimination technique; 
(b) accuracy of the OEP and EFP2 estimates of the CT energy with respect to the
CT/HF0 estimate. 
The structures of bi\hyp{}molecular complexes
were taken from the NCB31 database developed by the Truhlar's group:
the HB6/04 hydrogen bonding database 
(red),\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
the DI6/04 dipoleinteraction database 
(green),\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
the CT7/04 charge-transfer complex database 
(blue),\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
the WI7/05 weak interaction database 
(yellow),\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
and PPS5/05 the $\pi$-$\pi$ stacking database 
(black).\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
6-311++G(d,p) basis set was used for the CT energy calculations
and the single GDF scheme with the aug-cc-pVDZ auxiliary basis set
for the OEP calculations.
} 
\end{figure*}
%

@@

Even for the rather large auxiliary quadruple\hyp{}$\zeta$ basis set,
the OEP\hyp{}based model is roughly 9 times more efficient than evaluating the EFP2 CT energy.
Reducing the size of auxiliary basis set up to minimal basis from Table~\ref{t:3}
scales down the cost further by a factor of $1.9$ in water-methanol system,
and $2.4$ in water dimer system, reaching in total roughly 20\hyp{}fold
reduction in CPU time as compared to the EFP2 model, and more than three orders of magnitude
as compared to the OL model. Since the size of such a minimal basis
is very small as compared to the primary basis sets recommended for constructing
the EFP2 parameters, the additional superimposition costs of the auxiliary basis
should be negligible.
It is also worth mentioning here that, while using the usual 6-311++G(d,p) basis
in a role of auxiliary basis set 
gave already accurate results, the STO-3G basis as auxiliary space resulted
in CT energies of around $-8$ kcal/mol (data not shown). Therefore, the basis set optimization
described in Appendix~\ref{a:auxiliary-basis} was necessary for the minimal basis to be applicable
in the CT calculations using OEP technique, with OEP's generated by the EDF-2 scheme
developed in Section~\ref{sss:2.3.1.GDF-2}. 
%
%It is clear that the cost of the OEP\hyp{}based equations is much lesser than the cost of EFP2 method,
%and many orders of magnitude lesser than the original theory of Murrell et al.
%
\begin{figure}[t]
\includegraphics[width=0.46\textwidth]{fig-4.pdf}
\caption{\label{f:fig-4} {\bf Asymptotic dependence of the charge transfer energy
in the CT/HF0 formulation for methanol-warer complex with minimal OEP auxiliary basis.} 
Water molecule has been translated by $R$ from the starting geometry
optimized at HF/6-31+G(d,p) level,
along the vector specified in the insets.
The total interaction energy
is also shown for comparison in light grey color in this figure.
Interaction energies were obtained at the HF/6-311++G(d,p) level of theory.
In this figure, `OEP' denotes single GDF scheme with aug-cc-pVQZ-jkfit auxiliary basis set
whereas `OEP/mini` denotes double GDF scheme with the minimal auxiliary basis
given in Table~\ref{t:3}.
} 
\end{figure}
%

@@

\begin{acknowledgments}
This project is carried out under POLONEZ programme which has received funding from the European Union's
Horizon~2020 research and innovation programme under the Marie Sk{\l}odowska-Curie grant agreement 
No.~665778. This project is funded by National Science Centre, Poland 
(grant~no. 2016/23/P/ST4/01720) within the POLONEZ 3 fellowship.
We thank Professor Marcos Mandado from University of Vigo in Spain
for providing us with his benchmark calculations to validate our implementation
of the DDS method.
\end{acknowledgments}

@@

To fit the auxiliary DF basis for the treatment of the overlap\hyp{}like
OEP matrix elements with the operator $\hat{v}_{\rm eff}$, 
the following objective function is minimized
%
\begin{equation} \label{e:a1-obj}
 Z[\{\xi\}] = \sum_{\alpha i} \left[ 
     \tBraKet{\alpha}{\hat{v}_{\rm eff}}{i} - 
     \sum_{\xi}^{\rm DF} V_{\xi i} S_{\alpha \xi} 
    \right]^2 \;,
\end{equation}
%
where $\{\xi\}$ is the auxiliary basis set to optimize, whereas $\{\alpha\}$
is the `test' basis set used to probe the accuracy of the density fitting.
The forms of the %Pauli and 
charge\hyp{}transfer DF matrices $\tBraKet{\alpha}{\hat{v}_{\rm eff}}{i}$
can be directly derived 
from Eq.~\eqref{e:v-oep.ct}
%from Eqs.~\eqref{e:v-oep.rep} and \eqref{e:v-oep.ct}, respectively, 
by expanding the MO's in terms of atomic basis functions. The working formulae
are given below:
%
\begin{subequations} \label{e:a1-v.oep}
 \begin{align}
   \tBraKet{\alpha}{\hat{v}_{\rm eff}^{\rm Rep}}{i}
     &= \sum_{x}^{\rm At} W_{\alpha i}^{(x)} \nonumber   \\ 
        + \sum_{\beta\gamma\delta} &
           \left\{ 
             2 C_{\beta i} D_{\gamma\delta} - C_{\gamma i} D_{\beta \delta}
           \right\}
           \BraKet{\alpha\beta}{\gamma\delta} \;,\\
   \tBraKet{\alpha}{\hat{v}_{\rm eff}^{\rm CT}}{n} 
     &= \sum_{x}^{\rm At} W_{\alpha n}^{(x)} \nonumber   \\
       + \sum_{\beta\gamma\delta} &
           \left\{
             2 C_{\beta n} D_{\gamma\delta} - C_{\gamma n} D_{\beta \delta}
           \right\}
           \BraKet{\alpha\beta}{\gamma\delta}
    \;.
 \end{align}
\end{subequations}
%
In the above equations, ${\bf C}$ is the LCAO\hyp{}MO matrix whereas ${\bf D}$
is the one\hyp{}particle density matrix in AO basis.

@@

Overlap\hyp{}like OEP's are always associated with the effective one\hyp{}particle density
(or bond\hyp{}order) 
matrices and the $\lambda$ parameter from Eq.~\eqref{e:oep-operator} being either 1 or 0. 
Therefore, one can use this fact to compute the effective distributed multipole
moments. In this work, the cumulative atomic multipole moments (CAMM)
of Sokalski and Poirier with Mulliken partitioning of the AO space
were chosen because of their simplicity.
One of examples
of such OEP\hyp{}based models reported already in the literature is
the transition cumulative atomic multipole moments (TrCAMM) model
for estimation of the excitation energy transfer (EET) couplings in the F{\"o}rster
limit\cite{Blasiak.Maj.Cho.Gora.JCTC.2015}, 
or the vibrational solvatochromic distributed multipole moments
(SolCAMM) for the determination of the vibrational frequency shifts
of spatially localized IR probes.\cite{Blasiak.Lee.Cho.JCP.2013}
In general, having the OED and $\lambda$, the effective cumulative atomic charges
and dipole moments are given by
%
\begin{subequations} \label{e:a1-v.oep-camm}
 \begin{align} \label{e:a1-v.oep-camm.q}
  q_x^{\rm eff} &= \lambda Z_x - 
       \sum_{\alpha\in x}\sum_{\beta} 
        P_{\alpha\beta}^{\rm eff} \BraKet{\alpha}{\beta} \;, \\
  {\BM\upmu}_x^{\rm eff} &= 
       \sum_{\alpha\in x}\sum_{\beta} 
        P_{\alpha\beta}^{\rm eff}
   \left\{ 
    \BraKet{\alpha}{\beta} {\bf R}_x - \tBraKet{\alpha}{\bf r}{\beta} 
   \right\}
   \;,
 \end{align}
\end{subequations}
%
where ${\bf R}_x$ is the position vector of the $x$-th nucleus.
Note that $P_{\alpha\beta}^{\rm eff}$ can refer to the particular electron
spin, or their sum as a bond order matrix as well.
Equations for the effective distributed quadrupoles, octupoles and hexadecapoles
can be easily obtained by analogy from Ref.~\cite{Blasiak.Maj.Cho.Gora.JCTC.2015}

>>>>

The apparent difficulty in theoretically characterising the CT energy
in terms of the interacting molecular fragments
is indeed a challenge in developing modern force fields 
or \emph{ab initio} fragmentation methods\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}
for modeling structure
and dynamics in condensed phases.
This needs to be contrasted with the Coulombic electrostatics,
non\hyp{}Coulombic repulsion (i.e., due to Pauli exclusion principle), 
dispersion and induction for instance, 
which to a certain extent
can be well described by, respectively, 
the multipole moments of charge density distributions,\cite{Sokalski.Poirier.CPL.1983,
Etchebest.Lavery.Pullman.TheorChimActa.1982,Stone.JCTC.2005}
the van der Waals repulsive and attractive potentials, 
as well as the polarizability models, i.e., all being relatively simple and computationally
inexpensive to evaluate mathematical models. 
Due to this reason, the CT effects 
%while usually of significant and non\hyp{}negligible magnitude as compared 
%to the total interaction energy, 
are not explicitly included in any molecular mechanics force field developed up to date.
%such as OPLS, Amber, CHARMM, Gromos or even polarizale AMOEBA force field. 
The only way to
incorporate the CT effects in the condensed\hyp{}phase simulations,
apart from performing full QM electronic structure simulations,
is through the second generation of the effective fragment potential 
(EFP2) method,\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013,
   Nguyen.Pachter.Day.JCP.2014,
   Day.Jensen.Gordon.Webb.Stevens.Krauss.Garmer.Basch.Cohen.JCP.1996}
developed by Gordon group.\cite{Li.Gordon.Jensen.JCP.2006,Xu.Gordon.JCP.2013}
EFP2 is an \emph{ab initio} force field
derived on the grounds of the first\hyp{}principles perturbation theory
at the HF level and including the intermolecular dispersion effects
by the response theory. 
%based on the imaginary frequency\hyp{}dependent dipole\hyp{}dipole polarizabilities.
That is to say, the total intermolecular 
interaction potential, which can also be perceived as an approximation
to SAPT at HF level with added intermolecular dispersion correction, is approximated as 
%
\begin{equation}\label{e:u-efp2}
 U^{\rm EFP2} \approx U^{\rm Coul} + U^{\rm Ex-Rep} + U^{\rm Ind} + U^{\rm Disp} + U^{\rm CT} \;,
\end{equation}
%
where $U^{\rm Coul}$ is the Coulombic interaction energy of the unperturbed charge\hyp{}density
distributions of the monomers, approximated by the distributed multipole approximation,
$U^{\rm Ex-Rep}$ is the exchange\hyp{}repulsion energy originating from the Pauli exclusion
principle, that could be roughly understood as the Heitler\hyp{}London interaction
energy with $U^{\rm Coul}$ subtracted, $U^{\rm Ind}$ and $U^{\rm Disp}$
are the induction and dispersion energies obtained from the distributed polarizability
approximation, and finally $U^{\rm CT}$ is the CT energy, the focus of this work.

@@

                       \frac{ \varphi^*_\eta({\bf r}_1) \hat{v}^A \phi_i({\bf r}_2) } 

@@

Four complexes: 
(i) (H$_2$O)$_2$, 
(ii) H$_2$O--CH$_3$OH, 
(iii) H$_2$O--NH$_4^+$ and 
(iv) NO$_3^-$--NH$_4^+$,
were chosen as model systems to analyse the asymptotic dependence 
of CT/HF0 energy. 
The reference (zero\hyp{}displacement) geometries
were obtained by
energy\hyp{}optimizations at the HF/6-31+G(d,p) level,
as implemented in 
the {\sc Gaussian~16} quantum chemistry program package.\cite{Gaussian16}
Subsequently, 30 displaced geometries for each model complex
were obtained by translating one of the monomers along the vector 
co\hyp{}linear with the H-bond or N--N distance in the case of ammonium nitrate.
The reference structures as well as the translation vectors are
indicated in the insets of Figure~\ref{f:fig-1} and S1.
To perform statistical error analysis, structural databases
of bi\hyp{}molecular complexes in the non\hyp{}covalent
interactions database NCB31 developed by the Truhlar's 
group,\cite{Zhao.Schultz.Truhlar.JCTC.2006,
Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
as well as the BioFragment Database subset BBI for backbone\hyp{}backbone
interactions in proteins,\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017} 
as implemented in the {\sc Psi4}
program,\cite{Psi4.JCTC.2017}
were utilized.
In particular, the subsets from the NCB31 database were separately considered:
the HB6/04 hydrogen bonding database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
the DI6/04 dipoleinteraction database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
the CT7/04 charge-transfer complex database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
the WI7/05 weak interaction database,\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
and 
the PPS5/05 the $\pi$-$\pi$ stacking database.\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}

@@

%
\begin{figure*}[h]
\includegraphics[width=0.9\textwidth]{fig-2.pdf}
\caption{\label{f:fig-2} {\bf Performance of the OEP and EFP2 methods for the CT/HF0 charge transfer interaction energy
in small bi\hyp{}molecular neutral complexes.}
(a) accuracy of the ERI elimination technique; 
(b) accuracy of the OEP and EFP2 estimates of the CT energy with respect to the
CT/HF0 estimate. 
The structures of bi\hyp{}molecular complexes
were taken from the NCB31 database developed by the Truhlar's group:
the HB6/04 hydrogen bonding database 
(red),\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
the DI6/04 dipoleinteraction database 
(green),\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
the CT7/04 charge-transfer complex database 
(blue),\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
the WI7/05 weak interaction database 
(yellow),\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
and the PPS5/05 the $\pi$-$\pi$ stacking database 
(black).\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
6-311++G(d,p) basis set was used for the CT energy calculations
and the single GDF scheme with the aug-cc-pVDZ auxiliary basis set
for the OEP calculations.
} 
\end{figure*}
%

@@

Even for the rather large quadruple\hyp{}$\zeta$ auxiliary basis set,
the OEP\hyp{}based model is roughly 9 times more efficient than evaluating the EFP2 CT energy.
Reducing the size of auxiliary basis set up to minimal basis from Table~\ref{t:3}
scales down the cost further by a factor of $1.9$ in water-methanol system,
and $2.4$ in water dimer system, reaching in total roughly 20\hyp{}fold
reduction in CPU time as compared to the EFP2 model, and more than three orders of magnitude
as compared to the OL model. Since the size of such a minimal basis
is very small as compared to the primary basis sets recommended for constructing
the EFP2 parameters, the additional superimposition costs of the auxiliary basis
should be negligible.
It is also worth mentioning here that, while using the usual 6-311++G(d,p) basis
in a role of auxiliary basis set 
gave already accurate results, the STO-3G basis as auxiliary space resulted
in CT energies of around $-8$ kcal/mol (data not shown). Therefore, the basis set optimization
described in Appendix~\ref{a:auxiliary-basis} was necessary for the minimal basis to be applicable
in the CT calculations using OEP technique, with OEP's generated by the EDF-2 scheme
developed in Section~\ref{sss:2.3.1.GDF-2}. 
%
%It is clear that the cost of the OEP\hyp{}based equations is much lesser than the cost of EFP2 method,
%and many orders of magnitude lesser than the original theory of Murrell et al.
%
\begin{figure}[t]
\includegraphics[width=0.46\textwidth]{fig-4.pdf}
\caption{\label{f:fig-4} {\bf Asymptotic dependence of the charge transfer energy
in the CT/HF0 formulation for methanol-warer complex with minimal OEP auxiliary basis.} 
Water molecule has been translated by $R$ from the starting geometry
optimized at HF/6-31+G(d,p) level,
along the vector specified in the insets.
The total interaction energy
is also shown for comparison in light grey color in this figure.
Interaction energies were obtained at the HF/6-311++G(d,p) level of theory.
In this figure, `OEP' denotes single GDF scheme with aug-cc-pVQZ-jkfit auxiliary basis set
whereas `OEP/mini` denotes double GDF scheme with the minimal auxiliary basis
given in Table~\ref{t:3}.
} 
\end{figure}
%

@@

\begin{acknowledgments}
This project is carried out under POLONEZ programme which has received funding from the European Union's
Horizon~2020 research and innovation programme under the Marie Sk{\l}odowska-Curie grant agreement 
No.~665778. This project is funded by National Science Centre, Poland 
(grant~no. 2016/23/P/ST4/01720) within the POLONEZ 3 fellowship.
Wroclaw Centre for Networking and Supercomputing (WCSS) is acknowledged for
computational resources.
We thank Professor Marcos Mandado from University of Vigo in Spain
for providing us with his benchmark calculations to validate our implementation
of the DDS method.
\end{acknowledgments}

@@

To fit the auxiliary DF basis for the treatment of the overlap\hyp{}like
OEP matrix elements with the operator $\hat{v}_{\rm eff}$, 
the following objective function is minimized
%
\begin{equation} \label{e:a1-obj}
 Z[\{\xi\}] = \sum_{\alpha i} \left[ 
     \tBraKet{\alpha}{\hat{v}_{\rm eff}}{i} - 
     \sum_{\xi}^{\rm DF} V_{\xi i} S_{\alpha \xi} 
    \right]^2 \;,
\end{equation}
%
where $\{\xi\}$ is the auxiliary basis set to optimize, whereas $\{\alpha\}$
is the `test' basis set used to probe the accuracy of the density fitting.
The forms of the %Pauli and 
charge\hyp{}transfer DF matrices $\tBraKet{\alpha}{\hat{v}_{\rm eff}}{i}$
can be directly derived 
from Eq.~\eqref{e:v-oep.ct}
%from Eqs.~\eqref{e:v-oep.rep} and \eqref{e:v-oep.ct}, respectively, 
by expanding the MO's in terms of atomic basis functions. The working formula
is %given below:
%
\begin{multline} \label{e:a1-v.oep}
   \tBraKet{\alpha}{\hat{v}_{\rm eff}^{{\rm G}_1}}{n} 
     = \sum_{x}^{\rm At} W_{\alpha n}^{(x)}   \\
       + \sum_{\beta\gamma\delta} 
           \left\{
             2 C_{\beta n} D_{\gamma\delta} - C_{\gamma n} D_{\beta \delta}
           \right\}
           \BraKet{\alpha\beta}{\gamma\delta}
%    \;.
\end{multline}
%
for the OEP operator defined by 
%
\begin{equation}
 \sum_{\eta\in B}^{\rm DF} \hat{v}_{\rm eff}^{{\rm G}_1} \Ket{\eta} \equiv 
 \sum_{\eta\in B}^{\rm DF} V^B_{n\eta} \Ket{\eta} \;.
\end{equation}
%
In Eq.~\eqref{e:a1-v.oep}, ${\bf C}$ is the LCAO\hyp{}MO matrix whereas ${\bf D}$
is the one\hyp{}particle density matrix in AO basis.

@@

Overlap\hyp{}like OEP's are always associated with the effective one\hyp{}particle density
(or bond\hyp{}order) 
matrices and the $\lambda$ parameter from Eq.~\eqref{e:oep-operator} being either 1 or 0. 
Therefore, one can use this fact to compute the effective distributed multipole
moments. In this work, the cumulative atomic multipole moments (CAMM)
of Sokalski and Poirier with Mulliken partitioning of the AO space
were chosen because of their simplicity.
One of examples
of such OEP\hyp{}based models reported already in the literature is
the transition cumulative atomic multipole moments (TrCAMM) model
for estimation of the excitation energy transfer (EET) couplings in the F{\"o}rster
limit\cite{Blasiak.Maj.Cho.Gora.JCTC.2015}, 
or the vibrational solvatochromic distributed multipole moments
(SolCAMM) for the determination of the vibrational frequency shifts
of spatially localized IR probes.\cite{Blasiak.Lee.Cho.JCP.2013}
In general, having the OED and $\lambda$, the effective cumulative atomic charges
and dipole moments are given by
%
\begin{subequations} \label{e:a1-v.oep-camm}
 \begin{align} \label{e:a1-v.oep-camm.q}
  q_x^{\rm eff} &= \lambda Z_x - 
       \sum_{\alpha\in x}\sum_{\beta} 
        P_{\alpha\beta}^{\rm eff} \BraKet{\alpha}{\beta} \;, \\
  {\BM\upmu}_x^{\rm eff} &= 
       \sum_{\alpha\in x}\sum_{\beta} 
        P_{\alpha\beta}^{\rm eff}
   \left\{ 
    \BraKet{\alpha}{\beta} {\bf R}_x - \tBraKet{\alpha}{\hat{\bf r}}{\beta} 
   \right\}
   \;,
 \end{align}
\end{subequations}
%
where ${\bf R}_x$ is the position vector of the $x$-th nucleus.
Note that $P_{\alpha\beta}^{\rm eff}$ can refer to the particular electron
spin, or their sum as a bond order matrix as well.
Equations for the effective distributed quadrupoles, octupoles and hexadecapoles
can be easily obtained by analogy from Ref.~\cite{Blasiak.Maj.Cho.Gora.JCTC.2015}
